---
title: "Text_2_Recommations R Notebook"
output:
  html_document:
    df_print: paged
---

Hi,

## App details - Output tabs

Plan is to build a new app called **collaborative-filtering-app** which will have an 

1. *Overview* tab, 
2. an *Example datasets* tab, 
3. a *DTM descriptives tab*, 
4. a *IBCF recommendations* tab and finally 
5. a *UBCF recommendations* tab.

## App details - input UI

In terms of user input, the user: 

1. uploads DTM file (from either basic-textAn app or text-networks app), 
2. selects the focal.user (from dropdown) for whom prediction must be done, and
3. selects number of brand recommendations to show (using slider)

In this RMD, I demo some simple recommendation engine functionality (both user-based as well as item-based collaborative filtering) atop a simple DTM built from unstruc raw text data on brand preferences.

Behold.

P.S. I will assume that the raw corpus will be revised in the basic-text-an app to hard-code bigrams into the raw text.

Step 1 - DTM descriptives

```{r}
library(quanteda)
library(ggplot2)
library(magrittr)
library(text2vec)  # for sim2 func


# read-in DTM
#dtm <- read.csv(file.choose(), header=TRUE)
dtm <- read.csv("C:\\Users\\20052\\Dropbox\\teaching related\\MKTR\\Co2021\\session 10\\Co2021 session 10 data files\\Favourite OTT pgp21 _dtm.csv", header=TRUE)

## basic pre-proc
rownames(dtm) = dtm[,1]
dtm = dtm[,2:ncol(dtm)]

dim(dtm)  # DTM size must be shown in descriptives output
dtm[1:10, 1:10] # show this too in descriptives

## descriptive summary of DTM
a0 = colSums(dtm)
a1 = sort(a0, decreasing=TRUE, index.return=TRUE)
a2 = as.matrix(a0[a1$ix]) # Sorted freqs
token_freqs = data.frame(brands = rownames(a2), Freq = a2, Freq.Rank = seq(1, nrow(a2)))
rownames(token_freqs) = NULL
head(token_freqs, 10) # sorted freqs and ranks

```

## Step 2 - func to compute collaborative filtering (CF)s

```{r}

## setting input defaults
require(text2vec)
focal.user = "ox" # say. This comes from USER INPUT
n0 = 10 # top n0 unrated items user might like. again, USER INPUT

dtm2CF <- function(dtm, focal.user, n0){
  
  ## build token_freqs wala DF from DTM
  a0 = colSums(dtm)
  a1 = sort(a0, decreasing=TRUE, index.return=TRUE)
  a2 = as.matrix(a0[a1$ix]) # Sorted freqs
  token_freqs = data.frame(brands = rownames(a2), Freq = a2, Freq.Rank = seq(1, nrow(a2)))
  rownames(token_freqs) = NULL

  ## Finding similarities of all the item pairs.
  a3 = sim2(as.matrix(t(dtm)), method = "cosine", norm="l2") # item-item simil matrix
  a4 = sim2(as.matrix(dtm), method = "cosine", norm="l2") # user-user simil matrix
  
  
  ## IBCF: Generating the missing ratings in the table
  a5 = a4 %*% as.matrix(dtm) %*% a3 # 0.01s  single, simple step using matrix multiplication
  a5a = apply(a5, 1, function(x) {x/max(x)}) # 0 s!
  a5a = t(a5a)
  
  ## Find unrated items focal.user i0 might most like.
  i0 = which(rownames(a5a) == focal.user) # doc index num of focal.user
  a6 = a5a[i0, (dtm[i0,] == 0)] # retain only brands focal.user hasn't yet rated
  a7 = sort(a6, decreasing=TRUE, index.return=TRUE) # sort by predicted ratings
  output0_ibcf = data.frame(IBCF.Recommended.brands = names(a6[a7$ix[1:n0]]), 
                            Pref.Prob = as.matrix(round(a6[a7$ix[1:n0]], 3)))
  
  rownames(output0_ibcf) = NULL  
  test = dplyr::left_join(output0_ibcf, token_freqs, by = c("IBCF.Recommended.brands" = "brands"))
  output0_ibcf = test
  
  ## same for UBCF
  a8 = a4 %*% as.matrix(dtm)
  a9 = a4 %*% rep(1, nrow(a4))
  a10 = matrix(rep(as.vector(a9)), nrow(a4), ncol(a8))
  a11 = a8 / a10
  a12 = a11[i0, (dtm[i0,] == 0)] 
  a13 = sort(a12, decreasing=TRUE, index.return=TRUE) # 
  output0_ubcf = data.frame(UBCF.Recommended.brands = names(a12[a13$ix[1:n0]]), 
                            Pref.Prob = as.matrix(round(a12[a13$ix[1:n0]], 3)))
  
  rownames(output0_ubcf) = NULL
  test = dplyr::left_join(output0_ubcf, token_freqs, by = c("UBCF.Recommended.brands" = "brands"))
  output0_ubcf = test
  
  ## find most similar users to focal.user
  a4a = a4; diag(a4a) = 0   # set diag to 0
  a14 = sort(a4a[i0,], decreasing=TRUE, index.return=TRUE)
  a15 = a4a[i0, a14$ix]
  a16 = (a15 > 0) # names(a4a)[1:n0]
  most_simil_users_df = data.frame(most_similar_users = names(a15)[a16], 
                                   similarity_score = a15[a16])
  rownames(most_simil_users_df) = NULL
  
  return(list(output0_ibcf, output0_ubcf, most_simil_users_df)) 

  } # func ends

```

Let's test-drive above func
```{r}
## test-drive

system.time({ CF.list = dtm2CF(dtm, "groot", 12) })  # 0.06 s

ibcf.brands = CF.list[[1]]; ibcf.brands

ubcf.brands = CF.list[[2]]; ubcf.brands

simil.users = CF.list[[3]]; simil.users

```

Above two can be shown in their separate tabs. Have sortable HTML tables where necessary.

Any Qs, ask me.

Sudhir
