---
title: "RecommSys longDF Notebook"
output:
  html_document:
    df_print: paged
---

The Recommendation Systems wala app already exists that provides (a) IBCF, (b) UBCF and (c) User similarity scores, all based off of a adjacency matrix.

However, the adjacency matrix wala input requirement may just be a tad too restrictive. 

After all, in real life, most datasets we see are in classical *long form*, i.e. with users and items as columns with duplicates in either column allowed. 

Below is a function that takes such long form input and transforms it into *wide form* data - like the adjacency matrix - such that the present recomm sys app can read and process directly.

### Long Form Data example

Here's an example of long form data from Kaggle for a movies recomm system.

```{r data_load}

movies_reco_long = read.csv("movies_recomm_longform.csv")
dim(movies_reco_long)
head(movies_reco_long)
```

Notice the  dataset format and size.

All one needs for the recomm app are 2 (OK, 3) things, as below:

* a `user ID` column that tracks a user's prefs or purchases over time

* an `item ID` column that details what was preferred (movies, books, SKUs etc)

* a `rating` column (optional) in which users rated purchases. If this colm doesn;t exist, we'll create a colm of ones to track purchase incidence and use it instead.

For any long-form dataset input, user must specify the above 3 things via UI elements. 3 variable selection fields (drop-downs or click-choose) - one for each of the three. 

Note: In the `rating` field, give option of `NA` so that we can create a ones column instead if rating is unavailable.

Now we are ready to process the rest.

Behold.

```{r ui_inputs}
require(tidyverse)

df = movies_reco_long  # comes from fileInput in UI
user_id = df$user_id  # from UI, variable selection
item_id = df$item_id  # from UI, variable selection
rating = df$rating # from UI. Either rating colm exists or is NA

if (is.na(rating)) {
  rating = rep(1, length(user_id))
}

df0 = data.frame(user_id, item_id, rating)  # this goes to server.R as input
```

# Func to process longform

Time to code the func itself. Will use `dplyr` and `tidyr` functionality. Behold.

```{r func_def}

convert_longform <- function(df0){
  
  a1 = df0 %>% pivot_wider(id_cols = user_id, 
                names_from = item_id, 
                values_from = rating, 
                values_fill = 0)
  
  return(a1) } # func ends

```

Now can save the above and read-into recomm sys app as adjacency or wide-form matrix.

### Test-driving 

tidyverse is blazingly fast, takes <1 sec, even for the 100k obs in this logform dataset.

```{r test_drive}

longform_inp = movies_reco_long[,c("user_id", "item_id", "rating")]

system.time({
  adja_matrix = convert_longform(longform_inp)
})

dim(adja_matrix)
```

And now this `adja_matrix` object is ready to be processed as usual in the recomm sys app.

Note: Will need to add a check-box for user tospecify if input data is long-form or not. 

User can that way have option to directly upload adjacency matices (such as DTMs fom text apps).

Sudhir, June 2021